Using SIMD and SIMT vectorization to evaluate sparse chemical kinetic Jacobian matrices and thermochemical source terms

### 2018-08-09 paper.tex
fixed error, and replace revises with adds

In this article we
\begin{itemize}
 \item Derive and \add{reproducibly} verify a new Jacobian formulation that greatly increases sparsity\revisedelete{: \texttt{pyJac v2}};
 \item \revisedelete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
 \item \revise{Investigate the performance of SIMD\slash SIMT-vectorization for a wide of range chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}}; and finally}
 \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.
\end{itemize}
This work builds upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}\add{, using the new formulation, \texttt{pyJac v2}, to achieve these goals}.
To our knowledge is the first open-source, verified effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.

@@

\add{These efforts are made publicly available (see~\cref{A:availability}) via the open-source, high-performance, chemical kinetics code \texttt{pyJac}}.
The new molar-based formulation \delete{for \texttt{pyJac}} resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.
In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to \revisedelete{develop}\revise{incorporate} more sophisticated Jacobian approximations\delete{ based on the \texttt{pyJac} software} would be a worthwhile endeavor.

We also demonstrated source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
In addition to parallel OpenMP evaluation on the CPU, \revise{this work}\delete{the new version of \texttt{pyJac}} enabled the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
Deep vectorization is possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, but it yields no performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
Deep vectorization deserves further study with other platforms (e.g., CUDA).

>>>>

In this article we
\begin{itemize}
 \item Derive and \add{reproducibly} verify a new Jacobian formulation that greatly increases sparsity\delete{: \texttt{pyJac v2}};
 \item \delete{Enable}\add{Detail the implementation of} cross-platform SIMD\slash SIMT vectorization for CPUs, GPUs, and other accelerators;
 \item \add{Investigate the performance of SIMD\slash SIMT-vectorization for a wide of range chemical kinetic models, and compare with the previous version of our analytical chemical kinetic Jacobian code~\cite{Niemeyer:2016aa}; and finally}
 \item Discuss future extensions to this work as well as several promising directions for SIMD\slash SIMT vectorization in reactive-flow simulations.
\end{itemize}
This work builds upon our previous analytical chemical kinetic Jacobian code, \texttt{pyJac}~\cite{Niemeyer:2016aa}\add{, using the new formulation, \texttt{pyJac v2}, to achieve these goals}.
To our knowledge is the first open-source, verified effort that vectorizes the evaluation of chemical-kinetic source terms and Jacobian matrices for any chemical model on a wide selection of platforms.

@@

\add{These efforts are made publicly available (see~\cref{A:availability}) via the open-source, high-performance, chemical kinetics code \texttt{pyJac}}.
The new molar-based formulation \delete{for \texttt{pyJac}} resulted in highly sparse chemical kinetic Jacobians, and allows selection of either the constant-volume or constant-pressure approximation.
In addition, sparsity can be increased further by eliminating components associated with the bath gas, as discussed in \cref{S:sparsity}; this approximation is not a key feature of this work, and future efforts to \delete{develop}\add{incorporate} more sophisticated Jacobian approximations\delete{ based on the \texttt{pyJac} software} would be a worthwhile endeavor.

We also demonstrated source-term and Jacobian evaluation for a range of chemical kinetic models~\cite{Burke:2011fh,smith_gri-mech_30,Wang:2007,Sarathy:2013jr} and multiple CPUs\slash GPUs (\cref{t:cpus,t:gpus}).
In addition to parallel OpenMP evaluation on the CPU, \add{this work}\delete{the new version of \texttt{pyJac}} enabled the shallow-vectorized evaluation of the chemical-kinetic source terms and analytical Jacobian on both the CPU and GPU via OpenCL.
Deep vectorization is possible on the Portable OpenCL (POCL) platform~\cite{poclIJPP}, but it yields no performance benefit as POCL did not achieve vectorizations for any execution pattern studied.
Deep vectorization deserves further study with other platforms (e.g., CUDA).
